Enhanced proximity effect in zigzag-shaped Majorana Josephson junctions.

### 2019-10-30 paper/zigzag.tex
Minor grammar/interpunctuation fixes

\comment{The gap is small because of long trajectories.}
An important challenge in creating stable Majoranas is the appearance of a soft gap---a power law decay instead of an exponential decay of the density of states near zero energy.
In clean systems soft gap arises due to the reduction of the induced gap for states with the momentum directed along the junction~\cite{Gennes1963,Beenakker2005}.
From a semiclassical perspective, these momenta correspond to long paths through the semiconductor without interruption by the superconductor, shown in Fig.~\ref{fig:setup}(a).
These long trajectories have long flight times $\tau_\textrm{f} \approx L_\textrm{t} / v_F$ (see Fig.~\ref{fig:setup}), where $L_\textrm{t}$ is the trajectory length.
Equivalently, the Thouless energy of these trajectories $E_{\textrm{Th}}=\hbar / \tau_\textrm{f}$ is small, resulting in a small gap $E_\textrm{gap} \ll \Delta$.
This problem does not appear when the Fermi surface is small and the zero point motion dominates the transverse velocity, making a low filling of the bands a possible workaround\cite{Beenakker2005,Nijholt2016}.
However, tuning the system to a low chemical potential requires a precise knowledge of the band positions and makes the device more sensitive to disorder or microscopic inhomogeneities.
On the other hand, disorder scatters these long trajectories and introduces a cutoff on the scale of the mean free path~\cite{Golubov1988,Belzig1996,Pilgram2000} which Ref.~\onlinecite{Haim2019} proposes to use to improve Majorana properties; however, disorder is impossible to control to a required precision experimentally.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.9\columnwidth]{figures/zigzag.pdf}
\caption{The straight (top) and the zigzag (bottom) SNS junction.
The zigzag pattern has a peak-to-peak amplitude $z_y$ and a period $z_x$.
The yellow areas are superconductors with a phase difference of $\phi$ between the top and the bottom.
The middle area is the semiconductor of width $W$.
A magnetic field $B$ pointing in the $x$-direction causes a Zeeman splitting in the semiconductor
A trajectory traveling at a grazing angle (red curve) has a very long flight time $\tau_\textrm{f}$ and a very small induced gap $E_\textrm{gap} \ll \Delta$.
At the same time, the zigzag geometry limits the length of a trajectory therefore lowering $\tau_\textrm{f}$ and increasing $E_\textrm{gap}$.
\label{fig:setup}}
\end{figure}

@@

\comment{We calculate the topological phase diagram.}
In Fig.~\ref{fig:phasediagrams} we compare the phase diagrams of the straight and the zigzag junction.
We plot $E_\textrm{gap}$ as a function of magnetic field, chemical potential, $E_\textrm{gap}(B_x, \mu)$; and the superconducting phase difference $E_\textrm{gap}(B_x, \phi)$ for both a straight system [(c) and (e)] and a zigzag system [(d) and (f)].
Additionally, we plot the first $\SI{1300}{\nm}$ (one zigzag period) of the wave functions [(a) and (b)] at the optimal point in parameter space for the straight system.
For the straight system, we calculate $E_\textrm{gap}$ by performing a binary search in $E$ for the energy at which the propagating modes start to appear~\cite{Nijholt2016}.
Additionally, in Fig.~\ref{fig:phasediagrams}(c) we plot the phase boundaries obtained by solving a generalized eigenvalue problem~\cite{Nijholt2016}.
Due to the large size of the zigzag supercell, we are unable to apply these methods to zigzag geometries.
Instead, we calculate $E_\textrm{gap}$ by finding the absolute minimum of the spectrum $E_\textrm{gap}=\min{\left|E(k)\right|}$.
By both observing the gap closings and comparing to the topological phase diagram of the straight system, we then infer the topology of the zigzag system and verify this by calculating the Majorana wave function of a finite length zigzag.
As a further check, in App.~\ref{appendix:direction_B} we also compute $E_\textrm{gap}$ as function of the angle of magnetic field and observe that the zigzag device protects Majoranas from magnetic field misalignments.

@@

\comment{The zigzag geometry is a controllable tool in hardening the gap and decreasing Majorana size.}
The zigzag geometry increases the topological gap in the high density regime by more than an order of magnitude, as well as substantially reducing of Majorana size.
The improvements occur in a broad range of parameter values, moreover, even using $B_x$ optimal for the straight system in the high density regime (Fig.~\ref{fig:phasediagrams}), the Majorana size $\xi_\textrm{M}$ and $E_\textrm{gap}$ are still more than an order of magnitude better for the zigzag system.
We expect that the improvement of the device performance will significantly simplify the creation of Majorana devices and the detection of Majorana states.
The zigzag geometry offers a controllable way to remove long trajectories, making it easier to rely on than disorder~\cite{Haim2019}, that otherwise has a similar effect.

@@

\comment{Optimizing the geometry, more detailed modeling, and a simple analytical estimation are open questions for further research}
Our work is the first demonstration of the impact of the Majorana device geometry on its performance, and it opens a much harder question of finding the optimal geometry.
A promising approach to tackle this question would rely on constructing a quasiclassical model of the zigzag devices.
Finally, we have excluded several important physical effects, such as: disorder, electrostatics, the orbital effect of magnetic field, and the finite thickness of the sample.
While we expect these phenomena not to influence our qualitative findings, a more detailed simulation should provide better guidance to future experiments.

@@

\begin{figure}[!htb]
\centering
\includegraphics{figures/syst_snaking_magnetic_field.pdf}
\caption{
System where the direction of the magnetic field is parallel to the NS interface.
\label{fig:syst_snaking_magnetic_field}}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics{figures/snaking_magnetic_field.pdf}
\caption{
Comparison of the zigzag device's band structures with the magnetic field parallel to the NS interface [(a), (orange)] and a magnetic field in the $x$-direction (blue) for different values of $z_y$.
\label{fig:snaking_magnetic_field}}
\end{figure}

\comment{In a realistic device the magnetic field might not be inhomogeneous, but zigzag doesn't care.}

To verify that the change in the interface transparency alone is not sufficient to explain the increase of the topological gap, we have simulated the dispersion of a system with $z_x = z_y = \SI{80}{\nm} \approx v_\textrm{F}$ and $W=\SI{400}{\nm}$, so that the angle of the boundaries is sufficiently large, but the boundary modulation is much smaller than the width.
We have observed that $E_\textrm{gap}$ becomes $\SI{6.1}{\mu \eV}$, a two-fold increase compared to the straight system.

>>>>

\comment{The gap is small because of long trajectories.}
An important challenge in creating stable Majoranas is the appearance of a soft gap---a power law decay instead of an exponential decay of the density of states near zero energy.
In clean systems a soft gap arises due to the reduction of the induced gap for states with the momentum directed along the junction~\cite{Gennes1963,Beenakker2005}.
From a semiclassical perspective, these momenta correspond to long paths through the semiconductor without interruption by the superconductor, shown in Fig.~\ref{fig:setup}(a).
These long trajectories have long flight times $\tau_\textrm{f} \approx L_\textrm{t} / v_\textrm{F}$ (see Fig.~\ref{fig:setup}), where $L_\textrm{t}$ is the trajectory length and $v_\textrm{F}$ the Fermi velocity.
Equivalently, the Thouless energy of these trajectories $E_{\textrm{Th}}=\hbar / \tau_\textrm{f}$ is small, resulting in a small gap $E_\textrm{gap} \ll \Delta$.
This problem does not appear when the Fermi surface is small and the zero point motion dominates the transverse velocity, making a low filling of the bands a possible workaround\cite{Beenakker2005,Nijholt2016}.
However, tuning the system to a low chemical potential requires a precise knowledge of the band positions and makes the device more sensitive to disorder or microscopic inhomogeneities.
On the other hand, disorder scatters these long trajectories and introduces a cutoff on the scale of the mean free path~\cite{Golubov1988,Belzig1996,Pilgram2000} which Ref.~\onlinecite{Haim2019} proposes to use to improve Majorana properties; however, disorder is impossible to control to a required precision experimentally.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.9\columnwidth]{figures/zigzag.pdf}
\caption{The straight (top) and the zigzag (bottom) SNS junction.
The zigzag pattern has a peak-to-peak amplitude $z_y$ and a period $z_x$.
The yellow areas are superconductors with a phase difference of $\phi$ between the top and the bottom.
The middle area is the semiconductor of width $W$.
A magnetic field $B$ pointing in the $x$-direction causes a Zeeman splitting in the semiconductor.
A trajectory traveling at a grazing angle (red curve) has a very long flight time $\tau_\textrm{f}$ and a very small induced gap $E_\textrm{gap} \ll \Delta$.
At the same time, the zigzag geometry limits the length of a trajectory, therefore lowering $\tau_\textrm{f}$ and increasing $E_\textrm{gap}$.
\label{fig:setup}}
\end{figure}

@@

\comment{We calculate the topological phase diagram.}
In Fig.~\ref{fig:phasediagrams} we compare the phase diagrams of the straight and the zigzag junction.
We plot $E_\textrm{gap}$ as a function of magnetic field and chemical potential ($E_\textrm{gap}(B_x, \mu)$), and of magnetic field and superconducting phase difference ($E_\textrm{gap}(B_x, \phi)$) for both a straight system [(c) and (e)] and a zigzag system [(d) and (f)].
Additionally, we plot the first $\SI{1300}{\nm}$ (one zigzag period) of the wave functions [(a) and (b)] at the optimal point in parameter space for the straight system.
For the straight system, we calculate $E_\textrm{gap}$ by performing a binary search in $E$ for the energy at which the propagating modes start to appear~\cite{Nijholt2016}.
Additionally, in Fig.~\ref{fig:phasediagrams}(c) we plot the phase boundaries obtained by solving a generalized eigenvalue problem~\cite{Nijholt2016}.
Due to the large size of the zigzag supercell, we are unable to apply these methods to zigzag geometries.
Instead, we calculate $E_\textrm{gap}$ by finding the absolute minimum of the spectrum $E_\textrm{gap}=\min{\left|E(k)\right|}$.
By both observing the gap closings and comparing to the topological phase diagram of the straight system, we then infer the topology of the zigzag system and verify this by calculating the Majorana wave function of a finite length zigzag.
As a further check, in App.~\ref{appendix:direction_B} we also compute $E_\textrm{gap}$ as function of the angle of magnetic field and observe that the zigzag device protects Majoranas from magnetic field misalignments.

@@

\comment{The zigzag geometry is a controllable tool in hardening the gap and decreasing Majorana size.}
The zigzag geometry increases the topological gap in the high density regime by more than an order of magnitude, as well as substantially reduces Majorana size.
The improvements occur in a broad range of parameter values, moreover, even using $B_x$ optimal for the straight system in the high density regime (Fig.~\ref{fig:phasediagrams}), the Majorana size $\xi_\textrm{M}$ and $E_\textrm{gap}$ are still more than an order of magnitude better for the zigzag system.
We expect that the improvement of the device performance will significantly simplify the creation of Majorana devices and the detection of Majorana states.
The zigzag geometry offers a controllable way to remove long trajectories, making it easier to rely on than disorder~\cite{Haim2019}, that otherwise has a similar effect.

@@

\comment{Optimizing the geometry, more detailed modeling, and a simple analytical estimation are open questions for further research}
Our work is the first demonstration of the impact of the Majorana device geometry on its performance, and it opens a much harder question of finding the optimal geometry.
A promising approach to tackle this question would rely on constructing a quasiclassical model of the zigzag devices.
Finally, we have neglected several important physical effects, such as: disorder, electrostatics, the orbital effect of magnetic field, and the finite thickness of the sample.
While we expect these phenomena not to influence our qualitative findings, a more detailed simulation should provide better guidance to future experiments.

@@

\begin{figure}[!htb]
\centering
\includegraphics{figures/syst_snaking_magnetic_field.pdf}
\caption{
The magnetic field pattern emulating strong screening by the superconductors.
\label{fig:syst_snaking_magnetic_field}}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics{figures/snaking_magnetic_field.pdf}
\caption{
Comparison of the zigzag device's band structures with the magnetic field following the shape of the junction (orange) (see Fig.~\ref{fig:syst_snaking_magnetic_field}) and a magnetic field in the $x$-direction (blue) for different values of $z_y$.
\label{fig:snaking_magnetic_field}}
\end{figure}

\comment{In a realistic device the magnetic field might not be inhomogeneous, but zigzag doesn't care.}
The change in the interface transparency alone is not sufficient to explain the increase of the topological gap.
To verify this, we compute the dispersion of a system with $z_x = z_y = \SI{80}{\nm} \approx v_\textrm{F}$ and $W=\SI{400}{\nm}$, so that the angle of the boundaries is sufficiently large, but the boundary modulation is much smaller than the width.
We observe that $E_\textrm{gap}$ becomes $\SI{6.1}{\mu \eV}$, a two-fold increase compared to the straight system, not the order of magnitude increase resulting from the cutting off of long trajectories.
